---
title: "Wine Report"
author: "Abdulrahman Aleid"
date: '2022-06-15'
output: html_document

knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = "../"
    )
  })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


library(tidyverse)
library(skimr)
library(janitor)
library(here)
library(plotly)
library(GGally)
library(corrplot)
library(kableExtra)
library(viridis)

```
## Introduction
This report investigates the "Red Wine Quality" dataset which is  related to red variants of the Portuguese "Vinho Verde" wine...

## Dataset Exploration
Let's have a look at a summery of the dataset.

```{r, echo=FALSE}

# Import Red Wine Quality dataset.
wine <- read_csv(here("data", "winequality-red.csv"), show_col_types = FALSE)

# View data summary.
glimpse(wine)
skim(wine)

```
\
In the summary above we see a sample of each of the attributes of the dataset. We, also, see the distribution of each attribute represented by the standard deviation (sd), minimum (p0), lower [25th percentile] (p1), median [50th percentile] (p2), upper [75th percentile] (p3), and maximum [100th percentile] (p5) of each attribute.

___

The following table shows random 10 rows sample of the dataset.

```{r, echo=FALSE}

# Display random 10 rows of the dataset.
set.seed(1234)
kable(head(wine,10), "html") %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")

```
___

To further investigate the different variables, let's have a look at graph representations of them.

Starting by a distribution graph of the red wine quality rantings.


\
```{r, echo=FALSE}
invisible(dev.off())

suppressWarnings(min(wine))

# Plot a bar graph of the quality attribute.
plot_ly(data = wine, 
        x = ~quality,
        type = "histogram", 
        color = ~quality) %>%
  layout(xaxis = list(range=c(min(wine$quality),max(wine$quality))), 
         title = 'Distribution of Red Wine Quality Ratings')

```

The above graph suggests that the red wine quality ratings is symmetrically distributed.

If we consider the wine quality rating less than the median (`r median(wine$quality)` ranting) are poor wine, then lets see how many poor wine versus good wine we have in the dataset.

```{r, echo=FALSE}

# Plot number of good wine vs bad wine.
wine %>% 
  select(quality) %>% 
  mutate(goodQuality = quality >= 6) %>% 
ggplot(aes(goodQuality, fill = goodQuality)) +
  geom_bar() +
  geom_text(position = "stack", stat='count', aes(label=..count..), vjust = -0.5) +
  theme(plot.title=element_text(hjust=0.5)) +
  labs(x="Wine Quality",
       y="Observations",
       title = "Red Wine Quality: Good V.S. Bad")

```

From the binary red wine quality plot we observe that the number of bad quality wine is close to good once. This observation 

___

Next, let's look at the correlations between the different attributes.

```{r, echo=FALSE}

# Plot a heat-map to show the correlations between attributes.
nc=ncol(wine)
winedf <- wine[,1:12]
# winedf$quality <- as.integer(wine[,12])
correlations <- cor(winedf,method="pearson")
corrplot(correlations, number.cex = .9, method = "square", 
         hclust.method = "ward", order = "FPC",
         type = "full", tl.cex=0.8,tl.col = "black")

```

From this heat-map one can observe that the attributes that have the highest correlations with wine quality are alcohol, volatile acidity and sulphates. So let's have a closer look at them.

___

Let's take a look at the relationship between alcohol, sulphates and quality.

```{r, echo=FALSE}

wine %>%
  plot_ly(x = ~alcohol,  y = ~sulphates, color = ~quality, hoverinfo = 'text', colors = viridis(3),
 text = ~paste('Quality:', quality,
                        '<br>Alcohol:', alcohol,
                        '<br>sulphates:', sulphates)) %>%
  add_markers(opacity = 0.8) %>%
  layout(title = "3D Wine Quality",
         annotations=list(yref='paper',
                          xref="paper",
                          y=1.05,x=1.1, 
                          text="quality",
                          showarrow=F),
         scene = list(xaxis = list(title = 'Alcohol'),
                      yaxis = list(title = 'sulphates')))

```


## Conclusion
In conclusion, a good quality red wine has a strong correlations with alcohol, volatile acidity and sulphates. In order to have a good quality red wine volatile acidity and sulphates need to be at minimum unlike alcohol which need to be in a higher range.